---
title: "Ch 6 Exercise Examples"
author: "CW"
date: "April 24, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(fpp2)
library(seasonal)
library(gridExtra)
library(Mcomp)
library(stlplus)
```

# Ch 6 Exercise Examples

## Data

For these exercises we will use data from the M3 forecasting competition.  This particular one is Electric power, production by waterpower, total, monthly time series from 1982 to 1994, with the final 18 months being the test set.

```{r}
data <- (subset(M3,"INDUSTRY","monthly"))[3]
train <- data$N1878$x
test <- data$N1878$xx
```

## General Approach

These are a combination of questions 1, 5, 6, and 7/8.  Not all of each question is answered, but you should get the idea.

### Plotting

```{r}
autoplot(train)
ggsubseriesplot(train)
ggseasonplot(train)
```

This is clearly seasonal.  There may be a trend, and may be some cyclicality.  For classical decomposition purposes, multiplicative is probably best, because the seasonality in the middle is smaller than in the early and later years of the training set.

### Classical Decomposition

Here I'm going to show the automated and semi-automated way to do classical decomposition and seasonal adjustment.

```{r}
decomp <- decompose(train, type="multiplicative")
autoplot(decomp)

sadj.train <- seasadj(decomp)

sadj.train_fcast <- rwf(sadj.train, drift=TRUE, h=18)

autoplot(sadj.train_fcast) + autolayer(test)

reseason <- numeric(length=12)

     # You MUST create the vector you're going to populate, first.  If you don't know the length
     # of the ultimate vector, just initialize it as vector-name-here <- NULL.

reseason <- (decomp$seasonal[106:123])*(sadj.train_fcast$mean)

autoplot(sadj.train_fcast, PI=FALSE) + autolayer(reseason) + autolayer(test)

```

Suppose you want to recreate an original datapoint from deseasonalized data.

```{r}
obs_7_orig <- decomp$trend[7]*decomp$seasonal[7]*decomp$random[7]

obs_7_orig; train[7]
```

If you were using additive instead of multiplicative decomposition, you'd add the T,S, and R terms instead of multiplying them.

### Outliers

Consider the impact of outliers, and see how impacts vary depending on whether they are in the beginning, middle, or end of the series.

```{r}
x<- train
x[60] <- train[60] + 1000
plot1 <- x %>% decompose(., type = "multiplicative") %>% seasadj() %>% autoplot()

y<- train
y[15] <- train[15] + 1000
plot2 <- y %>% decompose(., type = "multiplicative") %>% seasadj() %>% autoplot()

z<- train
z[110] <- train[110] + 1000
plot3 <- z %>% decompose(., type = "multiplicative") %>% seasadj() %>% autoplot()

grid.arrange(plot1, plot2, plot3, ncol = 1)


autoplot((x %>% decompose(., type = "multiplicative") %>% seasadj()) - (y %>% decompose(., type = "multiplicative") %>% seasadj()))

autoplot((x %>% decompose(., type = "multiplicative") %>% seasadj()) - (z %>% decompose(., type = "multiplicative") %>% seasadj()))

autoplot((y %>% decompose(., type = "multiplicative") %>% seasadj()) - (z %>% decompose(., type = "multiplicative") %>% seasadj()))

```

## Forecasts with STL

```{r}
library(stlplus)

stlplus(train, s.window = 7) %>% plot_seasonal()
stlplus(train, s.window = 9) %>% plot_seasonal()
stlplus(train, s.window = 11) %>% plot_seasonal()

train.stl <- stl ( train, s.window = 9 )
plot(train.stl)

```

### Compare the STL results with those obtained using SEATS and X11.

```{r}
train.SEATS <- train %>% seas() %>% autoplot() + ggtitle("SEATS decomposition of `train`")
train.SEATS

train.X11 <- train %>% seas(x11 = "") %>% autoplot() + ggtitle("X11 decomposition of `train`")
train.X11

grid.arrange(train.SEATS,train.X11, nrow = 1)
```

### Plotting Seasonally Adjusted Data

Compute and plot the seasonally adjusted data.

```{r}
train_sadj <- seasadj ( stl ( train, s.window = 9, robust = TRUE))
autoplot(train_sadj) + autolayer(train)
```

### Use a naive method to produce forecasts of the seasonally adjusted data.

```{r}
train_sadj.naive <- rwf ( train_sadj, drift = TRUE)
autoplot(train_sadj.naive)
```

### Use stlf() to reseasonalize the results, giving forecasts for the original data.

```{r}
train.stlf <- stlf ( train, method = "naive" )
autoplot(train.stlf)
checkresiduals(train.stlf)
```

### STL vs Seasonal Naive

Compare forecasts from stlf() with those from snaive(), using a test set.

```{r}

train %>% stlf(s.window = 9, robust = TRUE) %>% accuracy(.,test)
train %>% snaive() %>% accuracy(.,test)
```
